On the entropy of isothermal dusty gas: application to protoplanetary disks

### 2016-08-01 paper.tex
regular update

\begin{abstract}
  Small dust particles embedded in gaseous protoplanetary disks are subject
  to strong dust-gas friction, causing them to almost follow the gas
  flow. This is similar to the conservation of entropy following an
  adiabatic fluid.  We show that in the strong drag
  limit that an isothermal dusty-gas indeed behaves like an  
  adiabatic pure gas, while finite dust-gas coupling may be
  interpreted as an effective 
  heating/cooling. We exploit this equivalence 
  to 
\begin{inparaenum}[1)]
\item
  define the entropy of isothermal dusty gas; and 
\item 
  generalize previous, well-known gaseous instabilities in
  protoplanetary disks to include dust dynamics. 
\end{inparaenum}
This includes the axisymmetric stability of strictly isothermal dusty
gas, from which we place quantitative constraints on the sharpness of 
dust edges that can persist in protoplanetary disks. We also extend
the vertical shear instability (VSI) in locally isothermal disks to include
dust. We find that  dust-loading reduces VSI growth rates. The presence of
strongly-coupled dust particles preferentially stabilize VSI modes with small radial
wavelengths, as well as `surface modes' concentrated near the
disk boundaries. We briefly discuss other applications of the 
equivalence between isothermal dusty gas and adiabatic hydrodynamics
to understand other dusty instabilities in protoplanetary disks. 
\end{abstract}

>>>>

\begin{abstract}
  Small dust particles embedded in gaseous protoplanetary disks are subject
  to strong dust-gas friction, causing them to almost follow the gas
  flow. This is similar to the conservation of entropy following an
  adiabatic fluid.  We show that in the strong drag
  limit that an isothermal dusty gas indeed behaves like an  
  adiabatic pure gas, and finite dust-gas coupling may be
  regarded as an effective 
  heating/cooling. We exploit this equivalence 
  to 
\begin{inparaenum}[1)]
\item
  define the entropy of isothermal dusty gas; and 
\item 
  generalize previous, well-known gaseous instabilities in
  protoplanetary disks to include dust dynamics. 
\end{inparaenum}
This includes the axisymmetric stability of strictly isothermal dusty
gas, from which we place quantitative constraints on the sharpness of 
dust edges that can persist in protoplanetary disks. We also extend
the vertical shear instability (VSI) in locally isothermal disks to include
dust. We find that  dust-loading reduces VSI growth rates. The presence of
strongly-coupled dust particles preferentially stabilize VSI modes with small radial
wavelengths, as well as `surface modes' concentrated near the
disk boundaries. We briefly discuss potential applications of the 
equivalence between isothermal dusty gas and standard hydrodynamics
to provide thermodynamical interpretations of  other dusty instabilities in protoplanetary disks. 
\end{abstract}
